---
title: "YouHeal Analysis"
subtitle: "Anthropometric Repeated Measure Reliability"
author: "Dom Grisafe"
date: "8/27/2018"
output: html_document
---

```{r}
#Initialize data
library(tidyverse)
library(haven)
library(psych)
library(ggplot2)
setwd("/Volumes/GoogleDrive/My Drive/20180823 YouHeal/Data")
YH_anthro_0 <- read.csv("Cohort 1 Anthropometric Data - Baseline.csv")
glimpse(YH_anthro_0)
YH_anthro_1 <- read.csv("Cohort 1 Anthropometric Data - Follow Up.csv")
glimpse(YH_anthro_1)
```

#Intraclass correlation coefficient (ICC)
A test of reliability of repeated measures.  
Read link below for a discussion of ICC.  
https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3670985/
```{r, echo = FALSE}
#Assess ICC of observations

#Tricep Skin Fold at Baseline
ICC_TF_0 <- ICC(YH_anthro_0[,2:4])

#Waist Circumference at Baseline
ICC_WC_0 <- ICC(YH_anthro_0[,7:9])

#Tricep Skin Fold at Follow-up
ICC_TF_1 <- ICC(YH_anthro_1[,2:4])

#Waist Circumference at Follow-up
ICC_WC_1 <- ICC(YH_anthro_1[,7:9])
```
#Summary Statistics and Intraclass correlation coefficient (ICC)
All Intraclass correlation coefficients are near 1.00, therefore the repeated measurements are consistent.  
If ICC is near 1.00, there is a high degree of correlation between measurements.  

***  

##Tricep Skin Fold at Baseline
```{r, echo=FALSE}
#build matrix
YH_stats_TF_0 <- matrix(NA, nrow=3, ncol=3)
rownames(YH_stats_TF_0) <- c("Measure 1", "Measure 2", "Measure 3")
colnames(YH_stats_TF_0) <- c("Mean","Std. Dev.", "N")

#assign means
YH_stats_TF_0[1,1] <- mean(YH_anthro_0$Tricep.Skinfold.1)
YH_stats_TF_0[2,1] <- mean(YH_anthro_0$Tricep.Skinfold.2)
YH_stats_TF_0[3,1] <- mean(YH_anthro_0$Tricep.Skinfold.3)

#assign standard deviations
YH_stats_TF_0[1,2] <- sd(YH_anthro_0$Tricep.Skinfold.1)
YH_stats_TF_0[2,2] <- sd(YH_anthro_0$Tricep.Skinfold.2)
YH_stats_TF_0[3,2] <- sd(YH_anthro_0$Tricep.Skinfold.3)

#assign sample size
YH_stats_TF_0[1,3] <- nrow(YH_anthro_0)
YH_stats_TF_0[2,3] <- nrow(YH_anthro_0)
YH_stats_TF_0[3,3] <- nrow(YH_anthro_0)

#print stats matrix
YH_stats_TF_0

#print ICC3
ICC_TF_0$results[3,]
#ICC3: A fixed set of k judges rate each target. 
#There is no generalization to a larger population of judges. 
#(MSB - MSE)/(MSB+ (nr-1)*MSE)
```

##Waist Circumference at Baseline
```{r, echo=FALSE}
#build matrix
YH_stats_WC_0 <- matrix(NA, nrow=3, ncol=3)
rownames(YH_stats_WC_0) <- c("Measure 1", "Measure 2", "Measure 3")
colnames(YH_stats_WC_0) <- c("Mean","Std. Dev.", "N")

#assign means
YH_stats_WC_0[1,1] <- mean(YH_anthro_0$Waist.Circumference.1)
YH_stats_WC_0[2,1] <- mean(YH_anthro_0$Waist.Circumference.2)
YH_stats_WC_0[3,1] <- mean(YH_anthro_0$Waist.Circumference.3)

#assign standard deviations
YH_stats_WC_0[1,2] <- sd(YH_anthro_0$Waist.Circumference.1)
YH_stats_WC_0[2,2] <- sd(YH_anthro_0$Waist.Circumference.2)
YH_stats_WC_0[3,2] <- sd(YH_anthro_0$Waist.Circumference.3)

#assign sample size
YH_stats_WC_0[1,3] <- nrow(YH_anthro_0)
YH_stats_WC_0[2,3] <- nrow(YH_anthro_0)
YH_stats_WC_0[3,3] <- nrow(YH_anthro_0)

#print stats matrix
YH_stats_WC_0

#print ICC3
ICC_WC_0$results[3,]
#ICC3: A fixed set of k judges rate each target. 
#There is no generalization to a larger population of judges. 
#(MSB - MSE)/(MSB+ (nr-1)*MSE)
```

##Tricep Skin Fold at Follow-up
```{r, echo=FALSE}
#build matrix
YH_stats_TF_1 <- matrix(NA, nrow=3, ncol=3)
rownames(YH_stats_TF_1) <- c("Measure 1", "Measure 2", "Measure 3")
colnames(YH_stats_TF_1) <- c("Mean","Std. Dev.", "N")

#assign means
YH_stats_TF_1[1,1] <- mean(YH_anthro_1$Tricep.Skinfold.1)
YH_stats_TF_1[2,1] <- mean(YH_anthro_1$Tricep.Skinfold.2)
YH_stats_TF_1[3,1] <- mean(YH_anthro_1$Tricep.Skinfold.3)

#assign standard deviations
YH_stats_TF_1[1,2] <- sd(YH_anthro_1$Tricep.Skinfold.1)
YH_stats_TF_1[2,2] <- sd(YH_anthro_1$Tricep.Skinfold.2)
YH_stats_TF_1[3,2] <- sd(YH_anthro_1$Tricep.Skinfold.3)

#assign sample size
YH_stats_TF_1[1,3] <- nrow(YH_anthro_1)
YH_stats_TF_1[2,3] <- nrow(YH_anthro_1)
YH_stats_TF_1[3,3] <- nrow(YH_anthro_1)

#print stats matrix
YH_stats_TF_1

#print ICC3
ICC_TF_1$results[3,]
#ICC3: A fixed set of k judges rate each target. 
#There is no generalization to a larger population of judges. 
#(MSB - MSE)/(MSB+ (nr-1)*MSE)
```

##Waist Circumference at Follow-up
```{r, echo=FALSE}
#build matrix
YH_stats_WC_1 <- matrix(NA, nrow=3, ncol=3)
rownames(YH_stats_WC_1) <- c("Measure 1", "Measure 2", "Measure 3")
colnames(YH_stats_WC_1) <- c("Mean","Std. Dev.", "N")

#assign means
YH_stats_WC_1[1,1] <- mean(YH_anthro_1$Waist.Circumference.1)
YH_stats_WC_1[2,1] <- mean(YH_anthro_1$Waist.Circumference.2)
YH_stats_WC_1[3,1] <- mean(YH_anthro_1$Waist.Circumference.3)

#assign standard deviations
YH_stats_WC_1[1,2] <- sd(YH_anthro_1$Waist.Circumference.1)
YH_stats_WC_1[2,2] <- sd(YH_anthro_1$Waist.Circumference.2)
YH_stats_WC_1[3,2] <- sd(YH_anthro_1$Waist.Circumference.3)

#assign sample size
YH_stats_WC_1[1,3] <- nrow(YH_anthro_1)
YH_stats_WC_1[2,3] <- nrow(YH_anthro_1)
YH_stats_WC_1[3,3] <- nrow(YH_anthro_1)

#print stats matrix
YH_stats_WC_1

#print ICC3
ICC_WC_1$results[3,]
#ICC3: A fixed set of k judges rate each target. 
#There is no generalization to a larger population of judges. 
#(MSB - MSE)/(MSB+ (nr-1)*MSE)
```

#Correlations betweeen average tricep skinfold (TS) and average waist circumference (WC) measurements
```{r, echo=FALSE}
baseCor <- cor(YH_anthro_0$Average.Tricep.Skinfold,YH_anthro_0$Average.Waist.Circumference)
follCor <- cor(YH_anthro_1$Average.Tricep.Skinfold,YH_anthro_1$Average.Waist.Circumference)
corr <- matrix(c(baseCor,follCor),nrow=1)
colnames(corr) <- c("Baseline","Follow Up")
rownames(corr) <- c("TS WC Correlations")
round(corr,3)
```

#Compare Tricep Skin Fold and Waist Circumference at Baseline and Follow-up
```{r, echo=FALSE}
#add time variable so know whether at baseline or follow-up
YH_anthro_0$time <- 0
YH_anthro_1$time <- 1

#merge datasets to singl dataset
YH_anthro_all <- rbind(YH_anthro_0, YH_anthro_1)
YH_anthro_all$time <- factor(YH_anthro_all$time,
                             levels = c(0,1),
                             labels = c("Baseline", "Follow-Up"))

#organize dataset by Survey ID variable
YH_anthro_all_sort <- YH_anthro_all[order(YH_anthro_all$Survey.ID),]

#boxplot of average tricep skin fold
ggplot(YH_anthro_all_sort, aes(x=time,group = time, y = Average.Tricep.Skinfold)) +
  geom_boxplot() +
  labs(title = "Comparing Average Tricep Skin Fold Measurements") +
  xlab("Time of Measurement") +
  ylab("Average Tricep Skin Fold Measurement (mm)")

#boxplot of average waist circumference
ggplot(YH_anthro_all_sort, aes(x=time,group = time, y = Average.Waist.Circumference)) +
  geom_boxplot() +
  labs(title = "Comparing Average Waist Circumference Measurements") +
  xlab("Time of Measurement") +
  ylab("Average Waist Circumference Measurement (cm)")
```

```{r, echo=FALSE}
#define which participants are obese

  #create empty vector of zeros and attach it to existing data
  obese <- rep(0, 26)
  YH_anthro_all_sort_obese <- cbind(YH_anthro_all_sort,obese)

  #vector with Survey.IDs of obese participants
  obeID <- c(700,878,831,341,080,899)
  
  #define obese variable as 1=="Yes"", if Survey.ID in the obeID vector
  for (i in 1:nrow(YH_anthro_all_sort_obese)){
    if (YH_anthro_all_sort_obese$Survey.ID[i] %in% obeID){
      YH_anthro_all_sort_obese$obese[i] = 1
    }
  }
  
  #label 0="No", 1="Yes" for obese variable
  YH_anthro_all_sort_obese$obese <- factor(YH_anthro_all_sort_obese$obese,levels=c(0,1),labels=c("No","Yes"))
  
#subset obese participants only
YH_stats_obese <- subset(YH_anthro_all_sort_obese, obese=="Yes")
```

#Boxplots restricted to obese participants only
```{r, echo=FALSE}
#make boxplots of variables
ggplot(YH_stats_obese, aes(x=time,group = time, y = Average.Tricep.Skinfold)) +
  geom_boxplot() +
  labs(title="Comparing Average Tricep Skin Fold Measurements", subtitle = "Restricted to obese participants (n=6)") +
  xlab("Time of Measurement") +
  ylab("Average Tricep Skin Fold Measurement (mm)")
  
#boxplot of average waist circumference
ggplot(YH_stats_obese, aes(x=time,group = time, y = Average.Waist.Circumference)) +
  geom_boxplot() +
  labs(title="Comparing Average Waist Circumference Measurements", subtitle = "Restricted to obese participants (n=6)") +
  xlab("Time of Measurement") +
  ylab("Average Waist Circumference Measurement (cm)")
```

```{r include=FALSE}

#Average inter-item correlation
#Rudimentary look at correlation between variables

#correlation between tricep fold measurements at baseline
print("Correlation between tricep fold measurements at baseline")
print("Measure 1 and 2")
cor(YH_anthro_0$Tricep.Skinfold.1,YH_anthro_0$Tricep.Skinfold.2)
print("Measure 1 and 3")
cor(YH_anthro_0$Tricep.Skinfold.1,YH_anthro_0$Tricep.Skinfold.3)
print("Measure 2 and 3")
cor(YH_anthro_0$Tricep.Skinfold.2,YH_anthro_0$Tricep.Skinfold.3)

#correlation between waist circumference measurements at baseline
print("Correlation between waist circumference measurements at baseline")
print("Measure 1 and 2")
cor(YH_anthro_0$Waist.Circumference.1,YH_anthro_0$Waist.Circumference.2)
print("Measure 1 and 3")
cor(YH_anthro_0$Waist.Circumference.1,YH_anthro_0$Waist.Circumference.3)
print("Measure 2 and 3")
cor(YH_anthro_0$Waist.Circumference.2,YH_anthro_0$Waist.Circumference.3)

#correlation between tricep fold measurements at baseline
print("Correlation between tricep fold measurements at Follow-up")
print("Measure 1 and 2")
cor(YH_anthro_1$Tricep.Skinfold.1,YH_anthro_1$Tricep.Skinfold.2)
print("Measure 1 and 3")
cor(YH_anthro_1$Tricep.Skinfold.1,YH_anthro_1$Tricep.Skinfold.3)
print("Measure 2 and 3")
cor(YH_anthro_1$Tricep.Skinfold.2,YH_anthro_1$Tricep.Skinfold.3)

#correlation between waist circumference measurements at baseline
print("Correlation between waist circumference measurements at Follow-up")
print("Measure 1 and 2")
cor(YH_anthro_1$Waist.Circumference.1,YH_anthro_1$Waist.Circumference.2)
print("Measure 1 and 3")
cor(YH_anthro_1$Waist.Circumference.1,YH_anthro_1$Waist.Circumference.3)
print("Measure 2 and 3")
cor(YH_anthro_1$Waist.Circumference.2,YH_anthro_1$Waist.Circumference.3)
```


```{r include=FALSE}
#Internal Consistency
#Methods taken from blog post below  
#https://www.r-bloggers.com/five-ways-to-calculate-internal-consistency/


#Install psych package

#install psych package to use cronbach alpha
#http://personality-project.org/r/psych/HowTo/getting_started.pdf

#install.packages("psych", dependencies = TRUE)
#install.packages(c("psych","GPArotation","sem"),dependencies=TRUE)
#install.packages("ctv")  #this downloads the task view package
#library(ctv) #this activates the ctv package
#install.views("Psychometrics") #one of the many Taskviews
```