---
title: "YouHeaL Cohort 1, Phases 1 & 2.1"
subtitle: "Data Management"
author: "Dom Grisafe"
date: "8/31/2018"
output: html_document
---

```{r, results = "hide"}
#Initialize libraries to be used
library(tidyverse)
  #tidyverse causes LaTeX error when outputting to PDF...change to "xelatex" from PDF options. Or use      HTML
library(haven)
library(psych)
library(ggplot2)
library(dplyr)
library(readr)
library(stringr)

#need to label variables
library(Hmisc)
```

#Anthropometric Data
Manage Anthropometric Data
```{r}
#set directory to "Data" folder where .CSV files are located
setwd("/Volumes/GoogleDrive/My Drive/20180823 YouHeal/Data")

#import data
YH_anthro_0 <- read.csv("Cohort 1 Anthropometric Data - Baseline.csv")
#glimpse(YH_anthro_0)
YH_anthro_1 <- read.csv("Cohort 1 Anthropometric Data - Follow Up.csv")
#glimpse(YH_anthro_1)

#delete blank columns titled "X"
YH_anthro_0$X <- NULL
YH_anthro_1$X <- NULL

#delete average measurment columns
YH_anthro_0$Average.Tricep.Skinfold <- NULL
YH_anthro_0$Average.Waist.Circumference <- NULL
YH_anthro_1$Average.Tricep.Skinfold <- NULL
YH_anthro_1$Average.Waist.Circumference <- NULL

#rename columns to easier things to work with
varNames_0 <- c("ID", "base_TSF_1", "base_TSF_2", "base_TSF_3", "base_WC_1", "base_WC_2", "base_WC_3")
colnames(YH_anthro_0) <- varNames_0

varNames_1 <- c("ID", "flup_TSF_1", "flup_TSF_2", "flup_TSF_3", "flup_WC_1", "flup_WC_2", "flup_WC_3")
colnames(YH_anthro_1) <- varNames_1

#merge datasets to single dataset
YH_anthro <- merge(YH_anthro_0, YH_anthro_1, by="ID")
```


#Sociodemographic and Survey Data
Manage Sociodemographic and Survey Data
```{r, results="hide"}
#set directory to "Data" folder where .CSV files are located
setwd("/Volumes/GoogleDrive/My Drive/20180823 YouHeal/Data")

#import data
#hella rows imported if do not limit to correct number of observations (n = 13 in phase 1)
YH_sociosurvey_0 <- read.csv("Cohort 1 Survey Data - Baseline.csv",nrows=13)
#glimpse(YH_sociosurvey_0)
YH_sociosurvey_1 <- read.csv("Cohort 1 Survey Data - Follow Up.csv",nrows=13)
#glimpse(YH_sociosurvey_1)

#rename columns to easier things to work with - baseline survey
varNames_0_survey <- c("ID", 
 "sex", "age", "language1", "language2", "english_prof", "race", "city", "zip",
 
 "base_Q7_YRBS_juice", "base_Q8_YRBS_fruit", "base_Q9_YRBS_salad",
 "base_Q10_YRBS_potato", "base_Q11_YRBS_carrot", "base_Q12_YRBS_otherveg", 
 "base_Q13_YRBS_soda", "base_Q14_YRBS_sportdrink", "base_Q15_YRBS_water", 
 "base_Q16_YRBS_milk", "base_Q17_YRBS_breakfast", "base_Q18_YRBS_allergy",
 "base_Q19_YRBS_physical", "base_Q20_YRBS_strength", "base_Q21_YRBS_tv",
 "base_Q22_YRBS_screentime", "base_Q23_YRBS_pe", "base_Q24_YRBS_sportsteam", 
 "base_Q25_YRBS_concussion", 
 
  "base_Q26_efficacy", "base_Q27_efficacy", "base_Q28_efficacy",
  "base_Q29_efficacy", "base_Q30_efficacy", "base_Q31_efficacy", 
  "base_Q32_efficacy", "base_Q33_efficacy", "base_Q34_efficacy", 
  "base_Q35_efficacy", 
 
  "base_Q36_empower", "base_Q37_empower", "base_Q38_empower",
  "base_Q39_empower", "base_Q40_empower", "base_Q41_empower",
  "base_Q42_empower", "base_Q43_empower", "base_Q44_empower",
  "base_Q45_empower", "base_Q46_empower", "base_Q47_empower"
)

#rename columns to easier things to work with - follow up survey
varNames_1_survey <- c("ID", 

 "flup_Q7_YRBS_juice", "flup_Q8_YRBS_fruit", "flup_Q9_YRBS_salad",
 "flup_Q10_YRBS_potato", "flup_Q11_YRBS_carrot", "flup_Q12_YRBS_otherveg", 
 "flup_Q13_YRBS_soda", "flup_Q14_YRBS_sportdrink", "flup_Q15_YRBS_water", 
 "flup_Q16_YRBS_milk", "flup_Q17_YRBS_breakfast", "flup_Q18_YRBS_allergy",
 "flup_Q19_YRBS_physical", "flup_Q20_YRBS_strength", "flup_Q21_YRBS_tv",
 "flup_Q22_YRBS_screentime", "flup_Q23_YRBS_pe", "flup_Q24_YRBS_sportsteam", 
 "flup_Q25_YRBS_concussion", 
 
  "flup_Q26_efficacy", "flup_Q27_efficacy", "flup_Q28_efficacy",
  "flup_Q29_efficacy", "flup_Q30_efficacy", "flup_Q31_efficacy", 
  "flup_Q32_efficacy", "flup_Q33_efficacy", "flup_Q34_efficacy", 
  "flup_Q35_efficacy", 
 
  "flup_Q36_empower", "flup_Q37_empower", "flup_Q38_empower",
  "flup_Q39_empower", "flup_Q40_empower", "flup_Q41_empower",
  "flup_Q42_empower", "flup_Q43_empower", "flup_Q44_empower",
  "flup_Q45_empower", "flup_Q46_empower", "flup_Q47_empower"
)

#assign column name vector to baseline and follow up surveys
colnames(YH_sociosurvey_0) <- varNames_0_survey
colnames(YH_sociosurvey_1) <- varNames_1_survey

#merge datasets to single dataset
YH_sociosurvey <- merge(YH_sociosurvey_0, YH_sociosurvey_1, by="ID")
#glimpse(YH_sociosurvey)
```

Merge Anthropometric and Survey Datasets
```{r, results="hide"}
#merge datasets to single dataset
YH_phase1 <- merge(YH_anthro, YH_sociosurvey, by="ID")
```


Derived Anthropometric Variables
```{r, results="hide"}
#average TSF and WC measurements at baseline and follow up
YH_phase1$base_TSF_avg <- (YH_phase1$base_TSF_1 + YH_phase1$base_TSF_2 + YH_phase1$base_TSF_3)/3
YH_phase1$flup_TSF_avg <- (YH_phase1$flup_TSF_1 + YH_phase1$flup_TSF_2 + YH_phase1$flup_TSF_3)/3
YH_phase1$base_WC_avg <- (YH_phase1$base_WC_1 + YH_phase1$base_WC_2 + YH_phase1$base_WC_3)/3
YH_phase1$flup_WC_avg <- (YH_phase1$flup_WC_1 + YH_phase1$flup_WC_2 + YH_phase1$flup_WC_3)/3


#cutoffs for obesity defined by the CDC >= 85th percentile for TSF and WC:
#Anthropometric Reference Data for Children and Adults: United States, 2007–2010
#https://www.google.com/search?q=Anthropometric+Reference+Data+for+Children+and+Adults%3A+United+States%2C+2007%E2%80%932010&oq=Anthropometric+Reference+Data+for+Children+and+Adults%3A+United+States%2C+2007%E2%80%932010&aqs=chrome..69i57.699j0j7&sourceid=chrome&ie=UTF-8

  #defining obesity by average tricep skinfold at baseline, by age and gender
  YH_phase1$obese_base_TSF <- 0
  for(i in 1:nrow(YH_phase1)){
    #for males, by age
    if(YH_phase1$sex[i] == "M"){
      if ((YH_phase1$age[i] == 15) & (YH_phase1$base_TSF_avg[i] >= 20.1)) {
        YH_phase1$obese_base_TSF[i] <- 1
      } else if ((YH_phase1$age[i] == 16) & (YH_phase1$base_TSF_avg[i] >= 18.3)) {
        YH_phase1$obese_base_TSF[i] <- 1
      } else if ((YH_phase1$age[i] == 17) & (YH_phase1$base_TSF_avg[i] >= 22.9)) {
        YH_phase1$obese_base_TSF[i] <- 1
      } 
    } else if(YH_phase1$sex[i] == "F"){
        #for females, by age
        if ((YH_phase1$age[i] == 15) & (YH_phase1$base_TSF_avg[i] >= 27.5)) {
          YH_phase1$obese_base_TSF[i] <- 1
        } else if ((YH_phase1$age[i] == 16) & (YH_phase1$base_TSF_avg[i] >= 26.0)) {
          YH_phase1$obese_base_TSF[i] <- 1
        } else if ((YH_phase1$age[i] == 17) & (YH_phase1$base_TSF_avg[i] >= 27.2)) {
          YH_phase1$obese_base_TSF[i] <- 1
        }
      }
    }

  #defining obesity by average waist circumference at baseline, by age and gender
  YH_phase1$obese_base_WC <- 0
  for(i in 1:nrow(YH_phase1)){
    #for males, by age
    if(YH_phase1$sex[i] == "M"){
      if ((YH_phase1$age[i] == 15) & (YH_phase1$base_WC_avg[i] >= 94.8)) {
        YH_phase1$obese_base_WC[i] <- 1
      } else if ((YH_phase1$age[i] == 16) & (YH_phase1$base_WC_avg[i] >= 99.3)) {
        YH_phase1$obese_base_WC[i] <- 1
      } else if ((YH_phase1$age[i] == 17) & (YH_phase1$base_WC_avg[i] >= 101.6)) {
        YH_phase1$obese_base_WC[i] <- 1
      } 
    } else if(YH_phase1$sex[i] == "F"){
        #for females, by age
        if ((YH_phase1$age[i] == 15) & (YH_phase1$base_WC_avg[i] >= 95.3)) {
          YH_phase1$obese_base_WC[i] <- 1
        } else if ((YH_phase1$age[i] == 16) & (YH_phase1$base_WC_avg[i] >= 92.2)) {
          YH_phase1$obese_base_WC[i] <- 1
        } else if ((YH_phase1$age[i] == 17) & (YH_phase1$base_WC_avg[i] >= 95.5)) {
          YH_phase1$obese_base_WC[i] <- 1
        }
      }
    }
  
    #label obesity by TSF
    YH_phase1$obese_base_TSF <- factor(YH_phase1$obese_base_TSF,
    levels = c(0,1),
    labels = c("Healthy","Obese"))
    
    #label obesity by WC
    YH_phase1$obese_base_WC <- factor(YH_phase1$obese_base_WC,
    levels = c(0,1),
    labels = c("Healthy","Obese"))
  
    #which participants are obese by tricep skinfold and waist circumference at baseline?
    print("Which participants are obese by TSF or WC > 85th percentile, by age and sex,             at baseline?")
    select(YH_phase1, ID, sex, age, base_TSF_avg, base_WC_avg, obese_base_TSF, obese_base_WC)
```

Label Anthropometric Derived Variables
```{r, results="hide"}
#Subject ID
label(YH_phase1$ID) <- "Subject ID"
  
#Tricep skin fold at baseline
label(YH_phase1$base_TSF_1) <- "Tricep Skin Fold 1 at Baseline (mm)"
label(YH_phase1$base_TSF_2) <- "Tricep Skin Fold 2 at Baseline (mm)"
label(YH_phase1$base_TSF_3) <- "Tricep Skin Fold 3 at Baseline (mm)"
  
#Average Waist circumference at baseline
label(YH_phase1$base_WC_1) <- "Waist Circumference 1 at Baseline (cm)"
label(YH_phase1$base_WC_2) <- "Waist Circumference 2 at Baseline (cm)"
label(YH_phase1$base_WC_3) <- "Waist Circumference 3 at Baseline (cm)"

#Tricep skin fold at follow up  
label(YH_phase1$flup_TSF_1) <- "Tricep Skin Fold 1 at Follow Up (mm)"
label(YH_phase1$flup_TSF_2) <- "Tricep Skin Fold 2 at Follow Up (mm)"
label(YH_phase1$flup_TSF_3) <- "Tricep Skin Fold 3 at Follow Up (mm)"
  
#Average Waist circumference at follow up
label(YH_phase1$flup_WC_1) <- "Waist Circumference 1 at Follow Up (cm)"
label(YH_phase1$flup_WC_2) <- "Waist Circumference 2 at Follow Up (cm)"
label(YH_phase1$flup_WC_3) <- "Waist Circumference 3 at Follow Up (cm)"

#Average Tricep skin fold at baseline and follow up
label(YH_phase1$base_TSF_avg) <- "Average Tricep Skin Fold at Baseline (mm)"
label(YH_phase1$flup_TSF_avg) <- "Average Tricep Skin Fold at Follow Up (mm)"

#Average Waist circumference at baseline and follow up
label(YH_phase1$base_WC_avg) <- "Average Waist Circumference at Baseline (cm)"
label(YH_phase1$flup_WC_avg) <- "Average Waist Circumference at Follow Up (cm)"

#Average Tricep skin fold at baseline and follow up
label(YH_phase1$obese_base_TSF) <- "Obese by TSF >85th percentile, by age and sex (CDC)"
label(YH_phase1$obese_base_WC) <- "Obese by WC >85th percentile, by age and sex (CDC)"
```

Scoring the Youth Risk Behavior Survey (YRBS) - Questions 72-83  
https://www.cdc.gov/healthyyouth/data/yrbs/data.htm  
```{r, results="hide"}
#Q7 - "During the past 7 days, how many times did you drink 100% fruit juices such as orange juice, apple juice, or grape juice? (Do not count punch, Kool-Aid, sports drinks, or other fruit-flavored drinks.)"
Q7_levels <- c(1,2,3,4,5,6,7)
Q7_labels <- c("Did not drink 100% fruit juice during the past 7 day", 
               "1 to 3 times during the past 7 day", 
               "4 to 6 times during the past 7 day",
               "1 time per day",
               "2 times per day",
               "3 times per day",
               "4 or more times per day")
YH_phase1$base_Q7_YRBS_juice <- factor(YH_phase1$base_Q7_YRBS_juice, levels = Q7_levels, labels = Q7_labels)
YH_phase1$flup_Q7_YRBS_juice <- factor(YH_phase1$flup_Q7_YRBS_juice, levels = Q7_levels, labels = Q7_labels)

#Q8 - "During the past 7 days, how many times did you eat fruit? (Do not count fruit juice.)"
Q8_levels <- c(1,2,3,4,5,6,7)
Q8_labels <- c("I did not eat fruit during the past 7 days", 
               "1 to 3 times during the past 7 days",
               "4 to 6 times during the past 7 days",
               "1 time per day",
               "2 times per day",
               "3 times per day",
               "4 or more times per day")
YH_phase1$base_Q8_YRBS_fruit <- factor(YH_phase1$base_Q8_YRBS_fruit, levels = Q8_levels, labels = Q8_labels)
YH_phase1$flup_Q8_YRBS_fruit <- factor(YH_phase1$flup_Q8_YRBS_fruit, levels = Q8_levels, labels = Q8_labels)

#Q9 - "During the past 7 days, how many times did you eat green salad?"
Q9_levels <- c(1,2,3,4,5,6,7)
Q9_labels <- c("I did not eat green salad during the past 7 days", 
               "1 to 3 times during the past 7 days",
               "4 to 6 times during the past 7 days",
               "1 time per day",
               "2 times per day",
               "3 times per day",
               "4 or more times per day")
YH_phase1$base_Q9_YRBS_salad <- factor(YH_phase1$base_Q9_YRBS_salad, levels = Q9_levels, labels = Q9_labels)
YH_phase1$flup_Q9_YRBS_salad <- factor(YH_phase1$flup_Q9_YRBS_salad, levels = Q9_levels, labels = Q9_labels)

#Q10 - "During the past 7 days, how many times did you eat potatoes? (Do not count french fries, fried potatoes, or potato chips.)"
Q10_levels <- c(1,2,3,4,5,6,7)
Q10_labels <- c("I did not eat potatoes during the past 7 days", 
               "1 to 3 times during the past 7 days",
               "4 to 6 times during the past 7 days",
               "1 time per day",
               "2 times per day",
               "3 times per day",
               "4 or more times per day")
YH_phase1$base_Q10_YRBS_potato <- factor(YH_phase1$base_Q10_YRBS_potato, levels = Q10_levels, labels = Q10_labels)
YH_phase1$flup_Q10_YRBS_potato <- factor(YH_phase1$flup_Q10_YRBS_potato, levels = Q10_levels, labels = Q10_labels)

#Q11 - "During the past 7 days, how many times did you eat carrots?"
Q11_levels <- c(1,2,3,4,5,6,7)
Q11_labels <- c("I did not eat carrots during the past 7 days", 
               "1 to 3 times during the past 7 days",
               "4 to 6 times during the past 7 days",
               "1 time per day",
               "2 times per day",
               "3 times per day",
               "4 or more times per day")
YH_phase1$base_Q11_YRBS_carrot <- factor(YH_phase1$base_Q11_YRBS_carrot, levels = Q11_levels, labels = Q11_labels)
YH_phase1$flup_Q11_YRBS_carrot <- factor(YH_phase1$flup_Q11_YRBS_carrot, levels = Q11_levels, labels = Q11_labels)

#Q12 - "During the past 7 days, how many times did you eat other vegetables? (Do not count green salad, potatoes, or carrots.)"
Q12_levels <- c(1,2,3,4,5,6,7)
Q12_labels <- c("I did not eat other vegetables during the past 7 days", 
               "1 to 3 times during the past 7 days",
               "4 to 6 times during the past 7 days",
               "1 time per day",
               "2 times per day",
               "3 times per day",
               "4 or more times per day")
YH_phase1$base_Q12_YRBS_otherveg <- factor(YH_phase1$base_Q12_YRBS_otherveg, levels = Q12_levels, labels = Q12_labels)
YH_phase1$flup_Q12_YRBS_otherveg <- factor(YH_phase1$flup_Q12_YRBS_otherveg, levels = Q12_levels, labels = Q12_labels)

#Q13 - "During the past 7 days, how many times did you drink a can, bottle, or glass of soda or pop, such as Coke, Pepsi, or Sprite? (Do not count diet soda or diet pop.)"
Q13_levels <- c(1,2,3,4,5,6,7)
Q13_labels <- c("I did not drink soda or pop during the past 7 days", 
               "1 to 3 times during the past 7 days",
               "4 to 6 times during the past 7 days",
               "1 time per day",
               "2 times per day",
               "3 times per day",
               "4 or more times per day")
YH_phase1$base_Q13_YRBS_soda <- factor(YH_phase1$base_Q13_YRBS_soda, levels = Q13_levels, labels = Q13_labels)
YH_phase1$flup_Q13_YRBS_soda <- factor(YH_phase1$flup_Q13_YRBS_soda, levels = Q13_levels, labels = Q13_labels)

#Q14 - "During the past 7 days, how many times did you drink a can, bottle, or glass of a sports drink such as Gatorade or Powerade? (Do not count low-calorie sports drinks such as Propel or G2.)"
Q14_levels <- c(1,2,3,4,5,6,7)
Q14_labels <- c("I did not drink sports drinks during the past 7 days", 
               "1 to 3 times during the past 7 days",
               "4 to 6 times during the past 7 days",
               "1 time per day",
               "2 times per day",
               "3 times per day",
               "4 or more times per day")
YH_phase1$base_Q14_YRBS_sportdrink <- factor(YH_phase1$base_Q14_YRBS_sportdrink, levels = Q14_levels, labels = Q14_labels)
YH_phase1$flup_Q14_YRBS_sportdrink <- factor(YH_phase1$flup_Q14_YRBS_sportdrink, levels = Q14_levels, labels = Q14_labels)

#Q15 - "During the past 7 days, how many times did you drink a bottle or glass of plain water? (Count tap, bottled, and unflavored sparkling water.)"
Q15_levels <- c(1,2,3,4,5,6,7)
Q15_labels <- c("I did not drink water during the past 7 days", 
               "1 to 3 times during the past 7 days",
               "4 to 6 times during the past 7 days",
               "1 time per day",
               "2 times per day",
               "3 times per day",
               "4 or more times per day")
YH_phase1$base_Q15_YRBS_water <- factor(YH_phase1$base_Q15_YRBS_water, levels = Q15_levels, labels = Q15_labels)
YH_phase1$flup_Q15_YRBS_water <- factor(YH_phase1$flup_Q15_YRBS_water, levels = Q15_levels, labels = Q15_labels)

#Q16 - "During the past 7 days, how many glasses of milk did you drink? (Count the milk you drank in a glass or cup, from a carton, or with cereal. Count the half pint of milk served at school as equal to one glass.)"
Q16_levels <- c(1,2,3,4,5,6,7)
Q16_labels <- c("I did not drink milk during the past 7 days", 
               "1 to 3 glasses during the past 7 days",
               "4 to 6 glasses during the past 7 days",
               "1 glass per day",
               "2 glasses per day",
               "3 glasses per day",
               "4 or more glasses per day")
YH_phase1$base_Q16_YRBS_milk <- factor(YH_phase1$base_Q16_YRBS_milk, levels = Q16_levels, labels = Q16_labels)
YH_phase1$flup_Q16_YRBS_milk <- factor(YH_phase1$flup_Q16_YRBS_milk, levels = Q16_levels, labels = Q16_labels)

#Q17 - "During the past 7 days, on how many days did you eat breakfast?"
Q17_levels <- c(1,2,3,4,5,6,7,8)
Q17_labels <- c("0 days", 
               "1 day",
               "2 days",
               "3 days",
               "4 days",
               "5 days",
               "6 days",
               "7 days")
YH_phase1$base_Q17_YRBS_breakfast <- factor(YH_phase1$base_Q17_YRBS_breakfast, levels = Q17_levels, labels = Q17_labels)
YH_phase1$flup_Q17_YRBS_breakfast <- factor(YH_phase1$flup_Q17_YRBS_breakfast, levels = Q17_levels, labels = Q17_labels)

#Q18 - "Are there any foods that you have to avoid because eating the food could cause an allergic reaction, like skin rashes, swelling, itching, vomiting, coughing, or trouble breathing?"
Q18_levels <- c(1,2,3)
Q18_labels <- c("Yes", 
               "No",
               "Not sure")
YH_phase1$base_Q18_YRBS_allergy <- factor(YH_phase1$base_Q18_YRBS_allergy, levels = Q18_levels, labels = Q18_labels)
YH_phase1$flup_Q18_YRBS_allergy <- factor(YH_phase1$flup_Q18_YRBS_allergy, levels = Q18_levels, labels = Q18_labels)

#Q19 - "During the past 7 days, on how many days were you physically active for a total of at least 60 minutes per day? (Add up all the time you spent in any kind of physical activity that increased your heart rate and made you breathe hard some of the time.)"
Q19_levels <- c(1,2,3,4,5,6,7,8)
Q19_labels <- c("0 days", 
               "1 day",
               "2 days",
               "3 days",
               "4 days",
               "5 days",
               "6 days",
               "7 days")
YH_phase1$base_Q19_YRBS_physical <- factor(YH_phase1$base_Q19_YRBS_physical, levels = Q19_levels, labels = Q19_labels)
YH_phase1$flup_Q19_YRBS_physical <- factor(YH_phase1$flup_Q19_YRBS_physical, levels = Q19_levels, labels = Q19_labels)

#Q20 - "During the past 7 days, on how many days did you do exercises to strengthen or tone your muscles, such as push-ups, sit-ups, or weight lifting?"
Q20_levels <- c(1,2,3,4,5,6,7,8)
Q20_labels <- c("0 days", 
               "1 day",
               "2 days",
               "3 days",
               "4 days",
               "5 days",
               "6 days",
               "7 days")
YH_phase1$base_Q20_YRBS_strength <- factor(YH_phase1$base_Q20_YRBS_strength, levels = Q20_levels, labels = Q20_labels)
YH_phase1$flup_Q20_YRBS_strength <- factor(YH_phase1$flup_Q20_YRBS_strength, levels = Q20_levels, labels = Q20_labels)

#Q21 - "On an average school day, how many hours do you watch TV?"
Q21_levels <- c(1,2,3,4,5,6,7)
Q21_labels <- c("I do not watch TV on an average school day", 
               "Less than 1 hour per day",
               "1 hour per day",
               "2 hours per day",
               "3 hours per day",
               "4 hours per day",
               "5 or more hours per day")
YH_phase1$base_Q21_YRBS_tv <- factor(YH_phase1$base_Q21_YRBS_tv, levels = Q21_levels, labels = Q21_labels)
YH_phase1$flup_Q21_YRBS_tv <- factor(YH_phase1$flup_Q21_YRBS_tv, levels = Q21_levels, labels = Q21_labels)

#Q22 - "On an average school day, how many hours do you play video or computer games or use a computer for something that is not school work? (Count time spent on things such as Xbox, PlayStation, an iPad or other tablet, a smartphone, texting, YouTube, Instagram, Facebook, or other social media.)"
Q22_levels <- c(1,2,3,4,5,6,7)
Q22_labels <- c("I do not play video or computer games or use a computer for something that is not school work", 
               "Less than 1 hour per day",
               "1 hour per day",
               "2 hours per day",
               "3 hours per day",
               "4 hours per day",
               "5 or more hours per day")
YH_phase1$base_Q22_YRBS_screentime <- factor(YH_phase1$base_Q22_YRBS_screentime, levels = Q22_levels, labels = Q22_labels)
YH_phase1$flup_Q22_YRBS_screentime <- factor(YH_phase1$flup_Q22_YRBS_screentime, levels = Q22_levels, labels = Q22_labels)

#Q23 - "In an average week when you are in school, on how many days do you go to physical education (PE) classes?"
Q23_levels <- c(1,2,3,4,5,6)
Q23_labels <- c("0 days", 
               "1 day",
               "2 days",
               "3 days",
               "4 days",
               "5 days")
YH_phase1$base_Q23_YRBS_pe <- factor(YH_phase1$base_Q23_YRBS_pe, levels = Q23_levels, labels = Q23_labels)
YH_phase1$flup_Q23_YRBS_pe <- factor(YH_phase1$flup_Q23_YRBS_pe, levels = Q23_levels, labels = Q23_labels)

#Q24 - "During the past 12 months, on how many sports teams did you play? (Count any teams run by your school or community groups.)"
Q24_levels <- c(1,2,3,4)
Q24_labels <- c("0 teams", 
               "1 team",
               "2 teams",
               "3 or more teams")
YH_phase1$base_Q24_YRBS_sportsteam <- factor(YH_phase1$base_Q24_YRBS_sportsteam, levels = Q24_levels, labels = Q24_labels)
YH_phase1$flup_Q24_YRBS_sportsteam <- factor(YH_phase1$flup_Q24_YRBS_sportsteam, levels = Q24_levels, labels = Q24_labels)

#Q25 - "During the past 12 months, how many times did you have a concussion from playing a sport or being physically active?"
Q25_levels <- c(1,2,3,4,5)
Q25_labels <- c("0 times", 
               "1 time",
               "2 times",
               "3 times",
               "4 or more times")
YH_phase1$base_Q25_YRBS_concussion <- factor(YH_phase1$base_Q25_YRBS_concussion, levels = Q25_levels, labels = Q25_labels)
YH_phase1$flup_Q25_YRBS_concussion <- factor(YH_phase1$flup_Q25_YRBS_concussion, levels = Q25_levels, labels = Q25_labels)

#look at YRBS at baseline
select(YH_phase1, "ID", 
 "base_Q7_YRBS_juice", "base_Q8_YRBS_fruit", "base_Q9_YRBS_salad",
 "base_Q10_YRBS_potato", "base_Q11_YRBS_carrot", "base_Q12_YRBS_otherveg", 
 "base_Q13_YRBS_soda", "base_Q14_YRBS_sportdrink", "base_Q15_YRBS_water", 
 "base_Q16_YRBS_milk", "base_Q17_YRBS_breakfast", "base_Q18_YRBS_allergy",
 "base_Q19_YRBS_physical", "base_Q20_YRBS_strength", "base_Q21_YRBS_tv",
 "base_Q22_YRBS_screentime", "base_Q23_YRBS_pe", "base_Q24_YRBS_sportsteam", 
 "base_Q25_YRBS_concussion"
)

#look at YRBS at follow up
select(YH_phase1, "ID",
 "flup_Q7_YRBS_juice", "flup_Q8_YRBS_fruit", "flup_Q9_YRBS_salad",
 "flup_Q10_YRBS_potato", "flup_Q11_YRBS_carrot", "flup_Q12_YRBS_otherveg", 
 "flup_Q13_YRBS_soda", "flup_Q14_YRBS_sportdrink", "flup_Q15_YRBS_water", 
 "flup_Q16_YRBS_milk", "flup_Q17_YRBS_breakfast", "flup_Q18_YRBS_allergy",
 "flup_Q19_YRBS_physical", "flup_Q20_YRBS_strength", "flup_Q21_YRBS_tv",
 "flup_Q22_YRBS_screentime", "flup_Q23_YRBS_pe", "flup_Q24_YRBS_sportsteam", 
 "flup_Q25_YRBS_concussion"
)
```

Scoring the General Self-Efficacy (GSE) Scale - Questions 26-35

```{r, results="hide"}
#Q26-Q35 - General Self-Efficacy (GSE) Scale 
Q26_Q35_efficacy_levels <- c(1,2,3,4)
Q26_Q35_efficacy_labels <- c("Not at all true",
               "Hardly true",
               "Moderately true",
               "Exactly true")

#Q26 - "I can always manage to solve difficult problems if I try hard enough"
YH_phase1$base_Q26_efficacy <- factor(YH_phase1$base_Q26_efficacy, levels = Q26_Q35_efficacy_levels, labels = Q26_Q35_efficacy_labels)
YH_phase1$flup_Q26_efficacy <- factor(YH_phase1$flup_Q26_efficacy, levels = Q26_Q35_efficacy_levels, labels = Q26_Q35_efficacy_labels)

#Q27 - "If someone opposes me, I can find the means and ways to get what I want."
YH_phase1$base_Q27_efficacy <- factor(YH_phase1$base_Q27_efficacy, levels = Q26_Q35_efficacy_levels, labels = Q26_Q35_efficacy_labels)
YH_phase1$flup_Q27_efficacy <- factor(YH_phase1$flup_Q27_efficacy, levels = Q26_Q35_efficacy_levels, labels = Q26_Q35_efficacy_labels)

#Q28 - "It is easy for me to stick to my aims and accomplish my goals."
YH_phase1$base_Q28_efficacy <- factor(YH_phase1$base_Q28_efficacy, levels = Q26_Q35_efficacy_levels, labels = Q26_Q35_efficacy_labels)
YH_phase1$flup_Q28_efficacy <- factor(YH_phase1$flup_Q28_efficacy, levels = Q26_Q35_efficacy_levels, labels = Q26_Q35_efficacy_labels)

#Q29 - "I am confident that I could deal efficiently with unexpected events."
YH_phase1$base_Q29_efficacy <- factor(YH_phase1$base_Q29_efficacy, levels = Q26_Q35_efficacy_levels, labels = Q26_Q35_efficacy_labels)
YH_phase1$flup_Q29_efficacy <- factor(YH_phase1$flup_Q29_efficacy, levels = Q26_Q35_efficacy_levels, labels = Q26_Q35_efficacy_labels)

#Q30 - "Thanks to my resourcefulness, I know how to handle unforeseen situations."
YH_phase1$base_Q30_efficacy <- factor(YH_phase1$base_Q30_efficacy, levels = Q26_Q35_efficacy_levels, labels = Q26_Q35_efficacy_labels)
YH_phase1$flup_Q30_efficacy <- factor(YH_phase1$flup_Q30_efficacy, levels = Q26_Q35_efficacy_levels, labels = Q26_Q35_efficacy_labels)

#Q31 - "I can solve most problems if I invest the necessary effort."
YH_phase1$base_Q31_efficacy <- factor(YH_phase1$base_Q31_efficacy, levels = Q26_Q35_efficacy_levels, labels = Q26_Q35_efficacy_labels)
YH_phase1$flup_Q31_efficacy <- factor(YH_phase1$flup_Q31_efficacy, levels = Q26_Q35_efficacy_levels, labels = Q26_Q35_efficacy_labels)

#Q32 - "I can remain calm when facing difficulties because I can rely on my coping abilities."
YH_phase1$base_Q32_efficacy <- factor(YH_phase1$base_Q32_efficacy, levels = Q26_Q35_efficacy_levels, labels = Q26_Q35_efficacy_labels)
YH_phase1$flup_Q32_efficacy <- factor(YH_phase1$flup_Q32_efficacy, levels = Q26_Q35_efficacy_levels, labels = Q26_Q35_efficacy_labels)

#Q33 - "When I am confronted with a problem, I can usually find several solutions."
YH_phase1$base_Q33_efficacy <- factor(YH_phase1$base_Q33_efficacy, levels = Q26_Q35_efficacy_levels, labels = Q26_Q35_efficacy_labels)
YH_phase1$flup_Q33_efficacy <- factor(YH_phase1$flup_Q33_efficacy, levels = Q26_Q35_efficacy_levels, labels = Q26_Q35_efficacy_labels)

#Q34 - "If I am in trouble, I can usually think of a solution"
YH_phase1$base_Q34_efficacy <- factor(YH_phase1$base_Q34_efficacy, levels = Q26_Q35_efficacy_levels, labels = Q26_Q35_efficacy_labels)
YH_phase1$flup_Q34_efficacy <- factor(YH_phase1$flup_Q34_efficacy, levels = Q26_Q35_efficacy_levels, labels = Q26_Q35_efficacy_labels)

#Q35 - "I can usually handle whatever comes my way."
YH_phase1$base_Q35_efficacy <- factor(YH_phase1$base_Q35_efficacy, levels = Q26_Q35_efficacy_levels, labels = Q26_Q35_efficacy_labels)
YH_phase1$flup_Q35_efficacy <- factor(YH_phase1$flup_Q35_efficacy, levels = Q26_Q35_efficacy_levels, labels = Q26_Q35_efficacy_labels)

#look at GSE at baseline
select(YH_phase1, "ID", 
 "base_Q26_efficacy","base_Q27_efficacy","base_Q28_efficacy","base_Q29_efficacy",
 "base_Q30_efficacy","base_Q31_efficacy","base_Q32_efficacy","base_Q33_efficacy",
 "base_Q34_efficacy","base_Q35_efficacy"
)

#look at GSE at baseline
select(YH_phase1, "ID", 
 "flup_Q26_efficacy","flup_Q27_efficacy","flup_Q28_efficacy","flup_Q29_efficacy",
 "flup_Q30_efficacy","flup_Q31_efficacy","flup_Q32_efficacy","flup_Q33_efficacy",
 "flup_Q34_efficacy","flup_Q35_efficacy"
)
```

Perceived Control Scale Items: Multiple Levels of Empowerment Indices
```{r, results="hide"}
#Q36-Q40 - Perceived Control Scale Items: Multiple Levels of Empowerment Indices
Q36_Q47_empowerment_levels <- c(1,2,3,4)
Q36_Q47_empowerment_labels <- c("Agree strongly",
               "Agree somewhat",
               "Disagree somewhat",
               "Disagree strongly")

#Q36 - "I can influence the decisions that this organization makes."
YH_phase1$base_Q36_empower <- factor(YH_phase1$base_Q36_empower, levels = Q36_Q47_empowerment_levels, labels = Q36_Q47_empowerment_labels)
YH_phase1$flup_Q36_empower <- factor(YH_phase1$flup_Q36_empower, levels = Q36_Q47_empowerment_levels, labels = Q36_Q47_empowerment_labels)

#Q37 - "This organization has influence over decisions that affect my life."
YH_phase1$base_Q37_empower <- factor(YH_phase1$base_Q37_empower, levels = Q36_Q47_empowerment_levels, labels = Q36_Q47_empowerment_labels)
YH_phase1$flup_Q37_empower <- factor(YH_phase1$flup_Q37_empower, levels = Q36_Q47_empowerment_levels, labels = Q36_Q47_empowerment_labels)

#Q38 - "This organization is effective in achieving its goals."
YH_phase1$base_Q38_empower <- factor(YH_phase1$base_Q38_empower, levels = Q36_Q47_empowerment_levels, labels = Q36_Q47_empowerment_labels)
YH_phase1$flup_Q38_empower <- factor(YH_phase1$flup_Q38_empower, levels = Q36_Q47_empowerment_levels, labels = Q36_Q47_empowerment_labels)

#Q39 - "This organization can influence decisions that affect the community."
YH_phase1$base_Q39_empower <- factor(YH_phase1$base_Q39_empower, levels = Q36_Q47_empowerment_levels, labels = Q36_Q47_empowerment_labels)
YH_phase1$flup_Q39_empower <- factor(YH_phase1$flup_Q39_empower, levels = Q36_Q47_empowerment_levels, labels = Q36_Q47_empowerment_labels)

#Q40 - "I am satisfied with the amount of influence I have over decisions that this organization makes."
YH_phase1$base_Q40_empower <- factor(YH_phase1$base_Q40_empower, levels = Q36_Q47_empowerment_levels, labels = Q36_Q47_empowerment_labels)
YH_phase1$flup_Q40_empower <- factor(YH_phase1$flup_Q40_empower, levels = Q36_Q47_empowerment_levels, labels = Q36_Q47_empowerment_labels)

#Q41 - "I have control over the decisions that affect my life."
YH_phase1$base_Q41_empower <- factor(YH_phase1$base_Q41_empower, levels = Q36_Q47_empowerment_levels, labels = Q36_Q47_empowerment_labels)
YH_phase1$flup_Q41_empower <- factor(YH_phase1$flup_Q41_empower, levels = Q36_Q47_empowerment_levels, labels = Q36_Q47_empowerment_labels)

#Q42 - "My community has influence over decisions that affect my life."
YH_phase1$base_Q42_empower <- factor(YH_phase1$base_Q42_empower, levels = Q36_Q47_empowerment_levels, labels = Q36_Q47_empowerment_labels)
YH_phase1$flup_Q42_empower <- factor(YH_phase1$flup_Q42_empower, levels = Q36_Q47_empowerment_levels, labels = Q36_Q47_empowerment_labels)

#Q43 - "I am satisfied with the amount of control I have over decisions that affect my life."
YH_phase1$base_Q43_empower <- factor(YH_phase1$base_Q43_empower, levels = Q36_Q47_empowerment_levels, labels = Q36_Q47_empowerment_labels)
YH_phase1$flup_Q43_empower <- factor(YH_phase1$flup_Q43_empower, levels = Q36_Q47_empowerment_levels, labels = Q36_Q47_empowerment_labels)

#Q44 - "I can influence decisions that affect my community."
YH_phase1$base_Q44_empower <- factor(YH_phase1$base_Q44_empower, levels = Q36_Q47_empowerment_levels, labels = Q36_Q47_empowerment_labels)
YH_phase1$flup_Q44_empower <- factor(YH_phase1$flup_Q44_empower, levels = Q36_Q47_empowerment_levels, labels = Q36_Q47_empowerment_labels)

#Q45 - "By working together, people in my community can influence decisions that affect the community."
YH_phase1$base_Q45_empower <- factor(YH_phase1$base_Q45_empower, levels = Q36_Q47_empowerment_levels, labels = Q36_Q47_empowerment_labels)
YH_phase1$flup_Q45_empower <- factor(YH_phase1$flup_Q45_empower, levels = Q36_Q47_empowerment_levels, labels = Q36_Q47_empowerment_labels)

#Q46 - "People in my community work together to influence decisions on the state or national level."
YH_phase1$base_Q46_empower <- factor(YH_phase1$base_Q46_empower, levels = Q36_Q47_empowerment_levels, labels = Q36_Q47_empowerment_labels)
YH_phase1$flup_Q46_empower <- factor(YH_phase1$flup_Q46_empower, levels = Q36_Q47_empowerment_levels, labels = Q36_Q47_empowerment_labels)

#Q47 - "I am satisfied with the amount of influence I have over decisions that affect my community."
YH_phase1$base_Q47_empower <- factor(YH_phase1$base_Q47_empower, levels = Q36_Q47_empowerment_levels, labels = Q36_Q47_empowerment_labels)
YH_phase1$flup_Q47_empower <- factor(YH_phase1$flup_Q47_empower, levels = Q36_Q47_empowerment_levels, labels = Q36_Q47_empowerment_labels)

#look at GSE at baseline
select(YH_phase1, "ID", 
 "base_Q36_empower", "base_Q37_empower", "base_Q38_empower", "base_Q39_empower", 
 "base_Q40_empower", "base_Q41_empower", "base_Q42_empower", "base_Q43_empower", 
 "base_Q44_empower", "base_Q45_empower", "base_Q46_empower", "base_Q47_empower"
)

#look at GSE at follow up
select(YH_phase1, "ID", 
 "flup_Q36_empower", "flup_Q37_empower", "flup_Q38_empower", "flup_Q39_empower", 
 "flup_Q40_empower", "flup_Q41_empower", "flup_Q42_empower", "flup_Q43_empower", 
 "flup_Q44_empower", "flup_Q45_empower", "flup_Q46_empower", "flup_Q47_empower"
)
```

#Scoring General Self Efficacy survey
The total score is calculated by finding the sum of the all items. For the GSE, the total score ranges between 10 and 40, with a higher score indicating more self-efficacy.  
  1 - Not at all true
  2 - Hardly true
  3 - Moderately true
  4 - Exactly true
```{r}
#look at baseline Self Efficacy items
glimpse(YH_phase1[,41:50])

#look at follow up Self Efficacy items
glimpse(YH_phase1[,82:91])

#get dataframe of Self Efficacy items, converting factors to doubles
#the factor values are actually the scoring weights, 
#so this is all that needs to be done

score_se <- function(itemcols){
  
  #get baseline Self Efficacy scores
  SE_base <- itemcols
  
  #make missing matrix the same size as the number of subjects:items in composite
  SE_dbl <- matrix(data = NA, nrow = nrow(SE_base), ncol = ncol(SE_base))
  
  #fill missing matrix with factor weights as doubles
  for (i in 1:nrow(SE_base)) {
    for (j in 1:ncol(SE_base)) {
      SE_dbl[i,j] <- as.double(SE_base[i,j])
    }
  }
  
  #return vector containing sum of all rows to make composite score
  return(rowSums(SE_dbl))
}

#look at composite Self Efficacy scores at baseline and follow up
base_Q26Q35_se_composite <- score_se(itemcols = YH_phase1[,41:50])
flup_Q26Q35_se_composite <- score_se(itemcols = YH_phase1[,82:91])

#matrix containing composite scores
se_composite <- matrix(c(base_Q26Q35_se_composite,flup_Q26Q35_se_composite), ncol = 2)
colnames(se_composite) <- c("base_Q26Q35_se_composite", "flup_Q26Q35_se_composite")

#add Self Efficacy composite scores matrix to complete dataset
YH_phase1_score <- cbind(YH_phase1, se_composite)
```

#Scoring Empowerment
Should be scored with the following coding:
  4 - Agree strongly
  3 - Agree somewhat
  2 - Disagree somewhat
  1 - Disagree strongly  

Note, this is reverse coding from how the data was entered 
```{r}
#look at baseline Self Efficacy items
glimpse(YH_phase1[,51:62])

#look at follow up Self Efficacy items
glimpse(YH_phase1[,92:103])

#note the factor values are inverted so greater value indicates less empowerment;
#invert factor values so greater value indicates greater empowerment

#get dataframe of Empowerment items, converting factors to doubles

score_em <- function(itemcols){
  
  #get baseline Empowerment scores
  EM_base <- itemcols
  
  #make missing matrix the same size as the number of subjects:items in composite
  EM_dbl <- matrix(data = NA, nrow = nrow(EM_base), ncol = ncol(EM_base))
  
  #invert factor values so greater value indicates greater empowerment
  
  #invert Empowerment factor weights so greater value is greater empowerment
  #fill missing matrix with modified factor weights, as doubles
  for (i in 1:nrow(EM_base)) {
    for (j in 1:ncol(EM_base)) {
      if(as.double(EM_base[i,j]) == 1){
        EM_dbl[i,j] <- 4
      } else if(as.double(EM_base[i,j]) == 2) {
        EM_dbl[i,j] <- 3
      } else if(as.double(EM_base[i,j]) == 3) {
        EM_dbl[i,j] <- 2
      } else if(as.double(EM_base[i,j]) == 4) {
        EM_dbl[i,j] <- 1
      } else {
        EM_base[i,j] <- NA
      }
    }
  }
  
  #return vector containing sum of all rows to make composite score
  return(rowSums(EM_dbl))
}

#look at overall Composite individual Empowerment scores at baseline and follow up
base_Q36Q47_em_composite <- score_em(itemcols = YH_phase1[,51:62])
flup_Q36Q47_em_composite <- score_em(itemcols = YH_phase1[,92:103])

#look at overall Individual Empowerment scores at baseline and follow up
base_Q41Q43_em_individual <- score_em(itemcols = matrix(c(YH_phase1[,56], YH_phase1[,58]), ncol = 2))
flup_Q41Q43_em_individual <- score_em(itemcols = matrix(c(YH_phase1[,97], YH_phase1[,99]), ncol = 2))

#look at overall Organizational Empowerment scores at baseline and follow up
base_Q36Q40_em_organizational <- score_em(itemcols = YH_phase1[,51:55])
flup_Q36Q40_em_organizational <- score_em(itemcols = YH_phase1[,92:96])

#look at overall Community Empowerment scores at baseline and follow up
base_Q42Q47_em_community <- score_em(itemcols = matrix(c(YH_phase1[,57], YH_phase1[,59], YH_phase1[,60], YH_phase1[,61], YH_phase1[,62]), ncol = 5))
flup_Q42Q47_em_community <- score_em(itemcols = matrix(c(YH_phase1[,98], YH_phase1[,100], YH_phase1[,101], YH_phase1[,102], YH_phase1[,103]), ncol = 5))

#matrix containing all 4 Empowerment composite scores at baseline and follow up
em_composite <- matrix(c(
    base_Q36Q47_em_composite, 
    base_Q41Q43_em_individual, 
    base_Q36Q40_em_organizational, 
    base_Q42Q47_em_community, 
    flup_Q36Q47_em_composite, 
    flup_Q41Q43_em_individual, 
    flup_Q36Q40_em_organizational, 
    flup_Q42Q47_em_community
  ), ncol = 8)

#assign column names
colnames(em_composite) <- c(
  "base_Q36Q47_em_composite", 
  "base_Q41Q43_em_individual", 
  "base_Q36Q40_em_organizational", 
  "base_Q42Q47_em_community", 
  "flup_Q36Q47_em_composite", 
  "flup_Q41Q43_em_individual", 
  "flup_Q36Q40_em_organizational", 
  "flup_Q42Q47_em_community"
  )

#add Empowerment composite scores matrix to complete dataset
YH_phase1_score <- cbind(YH_phase1_score, em_composite)
#glimpse(YH_phase1_score)
```



#Save Phase 1 Dataset
```{r}
#set save location
setwd("/Volumes/GoogleDrive/My Drive/20180823 YouHeal/Data/!Clean Datasets")

# Save phase 1 data file
#save(YH_phase1, file = "Phase1_Raw.RData")
#save(YH_phase1_score, file = "Phase1_Scored.RData")
```